% !TeX spellcheck = en_GB
% !TeX program = xelatex
% !TeX root = md2-user-handbook.tex

\subsection{Developing a Single App} 
\label{subsec:SingleAppDev}
In this section it is described how an application can be developed based on the current (March 2015) state of the MD2 DSL. To this end, this section is structured according to the different layers which constitute MD2 model. Aside from the conventional MVC layers -- model, view and controller -- this includes an additional layer, which enables and models the generation of workflows within and across apps and was established during the project seminar.

\subsubsection{Workflow} 
\label{subsubsec:Workflow}
The workflow layer is an additional abstraction on top of the controller layer. It thereby allows to specify the general course of action of one or more apps with few simple and well understandable model language constructs. Furthermore, this abtract workflow representation is intended to serve as a basis for communication with customers, e.g. for requirements engineering and collaborative app development.
In the workflow file of an MD2 model a workflow can be specified as a graph of workflow elements. Each workflow elements can fire one to many events. The respective event describes ... [to be continued]


\subsubsection{Model} 
\label{subsubsec:Model}

\subsubsection{View} 
\label{subsubsec:View}

\subsubsection{Controller} 
\label{subsubsec:Controller}


\subsection{Deploying a Single App}
\label{subsec:SingleAppDep}

\subsubsection{Backend} 
\label{subsubsec:Backend}

\subsubsection{map.apps} 
\label{subsubsec:mapapps}
