% !TeX spellcheck = en_US
% !TeX program = xelatex
% !TeX root = md2-user-handbook.tex

\MD is a framework for model-driven development of mobile business applications. It provides a domain specific language (DSL) for the specification of a textual model. Such a model can describe characteristics of a business scenario, including the communication between different apps and sequences of actions within each app, the views to be displayed and the data to be stored. A complete DSL model consists of four files, each representing a certain perspective -- model, view, controller as well as workflow -- on the application(s) to be generated.

In addition to the DSL, the \MD framework comprises a generator, which uses models specified in the DSL to generate source code. The code generation process creates source code for all apps that are specified in the model as well as a backend that handles their communication via web services as well as the communication with external applications. Originally, the \MD framework supported code generation for Android and IOS, but now includes a generator for a JavaScript-based web-platform called \mapapps. Updates for Android and IOS are intended in the future.

This handbook comprises a modeler's handbook and a developer's handbook. The modeler's handbook is targeted towards people who want to use the \MD framework to create models and generate code from them. It describes the constructs provided by the DSL and how to use them. Furthermore, it explains how the applications generated by the framework can be deployed. The developer's handbook aims to provide all information that is necessary for further development of and with the \MD framework. This includes the structure of the code generator for \mapapps as well as the structure and interaction of the generated and static code.
